---
title: "03_species_validation"
output: html_document
date: "2025-08-19"
---

---
#This script is to select audio files and annotations for validation purposes. Firstly, we will be randomly selecting 50% of the number of the detections, validating them on Raven Pro, and conducting logistic regression to arrive at the species and location-specific confidence thresholds.

##Loading required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)
```

##SPECIES- RVBU
#Reading the validation tables
```{r}
#Guwahati
GUW_RVBU_annotations <- read.delim ('E:/COVID-Soundscapes/Guwahati/GUW_validation/RVBU/GUW_RVBU_validation_table.txt', header = TRUE, sep = "\t")

#Delhi
DEL_RVBU_annotations <- read.delim ('E:/COVID-Soundscapes/Delhi/DEL_validation/RVBU/DEL_RVBU_validation_table.txt', header = TRUE, sep = "\t")

#Kodaikanal
KOD_RVBU_annotations <- read.delim ('E:/COVID-Soundscapes/Kodaikanal/KOD_validation/RVBU/KOD_RVBU_validation_table.txt', header = TRUE, sep = "\t")
```

#Extracting the confidence score values and adding it as a separate column
```{r}
#Guwahati
GUW_RVBU_annotations$conf_score <- substr(GUW_RVBU_annotations$Begin.File, 1, 5)

#Delhi
DEL_RVBU_annotations$conf_score <- substr(DEL_RVBU_annotations$Begin.File, 1, 5)

#Kodaikanal
KOD_RVBU_annotations$conf_score <- substr(KOD_RVBU_annotations$Begin.File, 1, 5)
```

#Fitting a logistic regression model and plotting the relationship between validation and confidence scores
```{r}
windowsFonts(A = windowsFont("Century Gothic")) #font for the plots

#Guwahati
GUW_RVBU_annotations$conf_score <- as.numeric (GUW_RVBU_annotations$conf_score) #reading confidence scores as numeric variables

GUW_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = GUW_RVBU_annotations) #fitting the logistic regression model with binary (0 or 1) outcome

prediction_values <- seq(0, 1, by = 0.1) #generating a sequence of hypothetical confidence scores

predictions <- predict(GUW_RVBU_LR, list(conf_score = prediction_values),  type = 'r') #obtaining the predicted probabilities of correct detections for given confidence scores

GUW_RVBU_plot <- plot(GUW_RVBU_annotations$valid ~ GUW_RVBU_annotations$conf_score, data= GUW_RVBU_annotations, xlim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(GUW_RVBU_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 2) #plotting the logistic regression curve to visualize the relationship between correct detections and confidence scores 

#The below code is to obtain the confidence score threshold. However, in the case of RVBU, most of our detections are below 0.5, so we might not be able to set a confidence score cut-off.
#p <- 0.97 #setting the probability of 0.95 (95% of detections)
#GUW_RVBU_threshold <- (log(p/(1-p))- GUW_RVBU_LR$coefficients[1]) /GUW_RVBU_LR$coefficients[2] 

#Delhi
DEL_RVBU_annotations$conf_score <- as.numeric (DEL_RVBU_annotations$conf_score) 

DEL_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = DEL_RVBU_annotations) 

prediction_values <- seq(0, 1, by = 0.1) 

predictions <- predict(DEL_RVBU_LR, list(conf_score = prediction_values),  type = 'r') 

DEL_RVBU_plot <- plot(DEL_RVBU_annotations$valid ~ DEL_RVBU_annotations$conf_score, data= DEL_RVBU_annotations, xlim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(DEL_RVBU_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 2)

#Kodaikanal
KOD_RVBU_annotations$conf_score <- as.numeric (KOD_RVBU_annotations$conf_score) 

KOD_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = KOD_RVBU_annotations) 

prediction_values <- seq(0, 1, by = 0.1) 

predictions <- predict(KOD_RVBU_LR, list(conf_score = prediction_values),  type = 'r') 

KOD_RVBU_plot <- plot(KOD_RVBU_annotations$valid ~ KOD_RVBU_annotations$conf_score, data= KOD_RVBU_annotations, xlim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(GUW_RVBU_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 3) 
```

##SPECIES- GRCO
#Reading the validation tables
```{r}
#Guwahati
GUW_GRCO_annotations <- read.delim ('E:/COVID-Soundscapes/Guwahati/GUW_validation/GRCO/GUW_GRCO_validation_table.txt', header = TRUE, sep = "\t")

#Delhi
DEL_GRCO_annotations <- read.delim ('E:/COVID-Soundscapes/Delhi/DEL_validation/GRCO/DEL_GRCO_validation_table.txt', header = TRUE, sep = "\t")

#Kodaikanal
KOD_GRCO_annotations <- read.delim ('E:/COVID-Soundscapes/Kodaikanal/KOD_validation/GRCO/KOD_GRCO_validation_table.txt', header = TRUE, sep = "\t")
```

#Extracting the confidence score values and adding it as a separate column
```{r}
#Guwahati
GUW_GRCO_annotations$conf_score <- substr(GUW_GRCO_annotations$Begin.File, 1, 5)

#Delhi
DEL_GRCO_annotations$conf_score <- substr(DEL_GRCO_annotations$Begin.File, 1, 5)

#Kodaikanal
KOD_GRCO_annotations$conf_score <- substr(KOD_GRCO_annotations$Begin.File, 1, 5)
```

#Fitting a logistic regression model and plotting the relationship between validation and confidence scores
```{r}
windowsFonts(A = windowsFont("Century Gothic")) 

#Guwahati
GUW_GRCO_annotations$conf_score <- as.numeric (GUW_GRCO_annotations$conf_score) 

GUW_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = GUW_GRCO_annotations) 

prediction_values <- seq(0, 1, by = 0.1) 

predictions <- predict(GUW_GRCO_LR, list(conf_score = prediction_values),  type = 'r') 

GUW_GRCO_plot <- plot(GUW_GRCO_annotations$valid ~ GUW_GRCO_annotations$conf_score, data= GUW_GRCO_annotations, xlim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(GUW_GRCO_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 2) 

p <- 0.95 
GUW_GRCO_threshold <- (log(p/(1-p))- GUW_GRCO_LR$coefficients[1]) /GUW_GRCO_LR$coefficients[2] 

#Delhi
DEL_GRCO_annotations$conf_score <- as.numeric (DEL_GRCO_annotations$conf_score) 

DEL_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = DEL_GRCO_annotations) 

prediction_values <- seq(0, 1, by = 0.1) 

predictions <- predict(DEL_GRCO_LR, list(conf_score = prediction_values),  type = 'r') 

DEL_GRCO_plot <- plot(DEL_GRCO_annotations$valid ~ DEL_GRCO_annotations$conf_score, data= DEL_GRCO_annotations, xlim=c(0,1), ylim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(DEL_GRCO_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 2)

#Kodaikanal
KOD_GRCO_annotations$conf_score <- as.numeric (KOD_GRCO_annotations$conf_score) 

KOD_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = KOD_GRCO_annotations) 

prediction_values <- seq(0, 1, by = 0.1) 

predictions <- predict(KOD_GRCO_LR, list(conf_score = prediction_values),  type = 'r') 

KOD_GRCO_plot <- plot(KOD_GRCO_annotations$valid ~ KOD_GRCO_annotations$conf_score, data= KOD_GRCO_annotations, xlim=c(0,1), pch=16, cex = 1.2, col=rgb(0,0,0,.3), xlab = "BirdNET Confidence Score", ylab = "True Positive (0/1)", family = 'A')  
curve(predict(GUW_GRCO_LR, data.frame(conf_score = x), type = "response"),  add = TRUE, col = "red", lwd = 3) 
```

