---
title: "03_species_validation"
output: html_document
date: "2025-08-19"
---

---
#This script is to select audio files and annotations for validation purposes. Firstly, we will be randomly selecting 50% of the number of the detections, validating them on Raven Pro, and conducting logistic regression to arrive at the species and location-specific confidence thresholds.

##Loading required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)
library(viridis)
library(sjPlot)
library(ggpubr)
library(stringr)
```

##SPECIES- RVBU
#Reading the validation tables
```{r}
#Guwahati
GUW_RVBU_annotations <- read.delim ('D:/COVID-Soundscapes/Guwahati/GUW_validation/RVBU/GUW_RVBU_validation_table.txt', header = TRUE, sep = "\t")

#Delhi
DEL_RVBU_annotations <- read.delim ('D:/COVID-Soundscapes/Delhi/DEL_validation/RVBU/DEL_RVBU_validation_table.txt', header = TRUE, sep = "\t")

#Kodaikanal
KOD_RVBU_annotations <- read.delim ('D:/COVID-Soundscapes/Kodaikanal/KOD_validation/RVBU/KOD_RVBU_validation_table.txt', header = TRUE, sep = "\t")
```

#Extracting the confidence score values and adding it as a separate column
```{r}
#Guwahati
GUW_RVBU_annotations$conf_score <- substr(GUW_RVBU_annotations$Begin.File, 1, 5)

#Delhi
DEL_RVBU_annotations$conf_score <- substr(DEL_RVBU_annotations$Begin.File, 1, 5)

#Kodaikanal
KOD_RVBU_annotations$conf_score <- substr(KOD_RVBU_annotations$Begin.File, 1, 5)
```

#Fitting a logistic regression model and plotting the relationship between validation and confidence scores
```{r}
windowsFonts(A = windowsFont("Century Gothic")) #font for the plots

#Guwahati
GUW_RVBU_annotations$conf_score <- as.numeric (GUW_RVBU_annotations$conf_score) #reading confidence scores as numeric variables

GUW_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = GUW_RVBU_annotations) #fitting the logistic regression model with binary (0 or 1) outcome

prediction_values <- seq(0, 1, .001) #generating a sequence of hypothetical confidence scores

predictions <- predict(GUW_RVBU_LR, list(conf_score = prediction_values),  type = 'r') #obtaining the predicted probabilities of correct detections for given confidence scores

#The below code is to obtain the confidence score threshold. 
GUW_RVBU_90 <- (log(.90 / (1 - .90)) - GUW_RVBU_LR$coefficients[1]) / GUW_RVBU_LR$coefficients[2] #confidence scorees beyond 0.248 will yield 90% true detections
GUW_RVBU_95 <- (log(.95 / (1 - .95)) - GUW_RVBU_LR$coefficients[1]) / GUW_RVBU_LR$coefficients[2] #confidence scorees beyond 0.596 will yield 95% true detections
GUW_RVBU_99 <- (log(.99 / (1 - .99)) - GUW_RVBU_LR$coefficients[1]) / GUW_RVBU_LR$coefficients[2] #this probability of detections is unattainable

#get viridis colors (colour-blind friendly)
viz_colors <- viridis(5)
curve_color <- viz_colors[1]
threshold_colors <- viz_colors[c(2, 3, 4)]

#rounding threshold values to 3 decimal places for labels
GUW_RVBU_90_label <- round(GUW_RVBU_90, 3)
GUW_RVBU_95_label <- round(GUW_RVBU_95, 3)
GUW_RVBU_99_label <- round(GUW_RVBU_99, 3)

GUW_RVBU_LR_plot <- ggplot(GUW_RVBU_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
    geom_vline(xintercept = GUW_RVBU_90, color = threshold_colors[1], size = 1) +
    geom_vline(xintercept = GUW_RVBU_95, color = threshold_colors[2], size = 1) +
    geom_vline(xintercept = GUW_RVBU_99, color = threshold_colors[3], size = 1) +
    # add threshold labels with offset
    annotate("text",
      x = pmin(pmax(GUW_RVBU_90, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("90%:", GUW_RVBU_90_label),
      color = threshold_colors[1],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    annotate("text",
      x = pmin(pmax(GUW_RVBU_95, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("95%:", GUW_RVBU_95_label),
      color = threshold_colors[2],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    #annotate("text", x = pmin(pmax(GUW_RVBU_99, 0.05), 0.95) + 0.02, y = 0.5, label = paste("99%:", GUW_RVBU_99_label), color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) + #omitting this because the corresponding probability is unattainable
    labs(
      x = "BirdNET Confidence Scores",
      y = "Prediction (0/1) "
    ) +
    theme_bw() +
    theme(text = element_text(family = "A"),
      axis.title = element_text(family = "A", size = 16, face = 'bold'),
      axis.text = element_text(family = "A", size = 12)
    ) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))

#Delhi
DEL_RVBU_annotations$conf_score <- as.numeric (DEL_RVBU_annotations$conf_score) 

DEL_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = DEL_RVBU_annotations) 

prediction_values <- seq(0, 1, .001) 

predictions <- predict(DEL_RVBU_LR, list(conf_score = prediction_values),  type = 'r') 

DEL_RVBU_90 <- (log(.90 / (1 - .90)) - DEL_RVBU_LR$coefficients[1]) / DEL_RVBU_LR$coefficients[2]
DEL_RVBU_95 <- (log(.95 / (1 - .95)) - DEL_RVBU_LR$coefficients[1]) / DEL_RVBU_LR$coefficients[2]
DEL_RVBU_99 <- (log(.99 / (1 - .99)) - DEL_RVBU_LR$coefficients[1]) / DEL_RVBU_LR$coefficients[2]

DEL_RVBU_90_label <- round(DEL_RVBU_90, 3)
DEL_RVBU_95_label <- round(DEL_RVBU_95, 3)
DEL_RVBU_99_label <- round(DEL_RVBU_99, 3)


DEL_RVBU_LR_plot <- ggplot(DEL_RVBU_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
   geom_vline(xintercept = DEL_RVBU_90, color = threshold_colors[1], size = 1) +
    geom_vline(xintercept = DEL_RVBU_95, color = threshold_colors[2], size = 1) +
    geom_vline(xintercept = DEL_RVBU_99, color = threshold_colors[3], size = 1) +
    # add threshold labels with offset
    annotate("text", x = pmin(pmax(DEL_RVBU_90, 0.05), 0.95) + 0.02, y = 0.5, label = paste("90%:", DEL_RVBU_90_label), color = threshold_colors[1], family = "A", angle = 90, hjust = 0.5, size = 5) +
    annotate("text", x = pmin(pmax(DEL_RVBU_95, 0.05), 0.95) + 0.02, y = 0.5, label = paste("95%:", DEL_RVBU_95_label), color = threshold_colors[2], family = "A", angle = 90, hjust = 0.5, size = 5) +
    annotate("text", x = pmin(pmax(DEL_RVBU_99, 0.05), 0.95) + 0.02, y = 0.5, label = paste("99%:", DEL_RVBU_99_label), color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) + #omitting because the corresponding probability is unattainable for all % cut-offs
    labs(
      x = "BirdNET Confidence Scores",
      y = "Prediction (0/1) "
    ) +
    theme_bw() +
    theme(text = element_text(family = "A"),
      axis.title = element_text(family = "A", size = 16, face = 'bold'),
      axis.text = element_text(family = "A", size = 12)
    ) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))

#Kodaikanal
KOD_RVBU_annotations$conf_score <- as.numeric (KOD_RVBU_annotations$conf_score) 

KOD_RVBU_LR <- glm(valid ~ conf_score, family = 'binomial', data = KOD_RVBU_annotations) 

prediction_values <- seq(0, 1, .001) 

predictions <- predict(KOD_RVBU_LR, list(conf_score = prediction_values),  type = 'r') 

KOD_RVBU_90 <- (log(.90 / (1 - .90)) - KOD_RVBU_LR$coefficients[1]) / KOD_RVBU_LR$coefficients[2] #confidence scores beyond 0.629 will yield 90% true detections
KOD_RVBU_95 <- (log(.95 / (1 - .95)) - KOD_RVBU_LR$coefficients[1]) / KOD_RVBU_LR$coefficients[2] #confidence scorees beyond 0.729 will yield 95% true detections
KOD_RVBU_99 <- (log(.99 / (1 - .99)) - KOD_RVBU_LR$coefficients[1]) / KOD_RVBU_LR$coefficients[2] #confidence scorees beyond 0.952 will yield 99% true detections

KOD_RVBU_90_label <- round(KOD_RVBU_90, 3)
KOD_RVBU_95_label <- round(KOD_RVBU_95, 3)
KOD_RVBU_99_label <- round(KOD_RVBU_99, 3)

KOD_RVBU_LR_plot <- ggplot(KOD_RVBU_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
    geom_vline(xintercept = KOD_RVBU_90, color = threshold_colors[1], size = 1) +
    geom_vline(xintercept = KOD_RVBU_95, color = threshold_colors[2], size = 1) +
    geom_vline(xintercept = KOD_RVBU_99, color = threshold_colors[3], size = 1) +
    annotate("text",
      x = pmin(pmax(KOD_RVBU_90, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("90%:", KOD_RVBU_90_label),
      color = threshold_colors[1],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    annotate("text",
      x = pmin(pmax(KOD_RVBU_95, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("95%:", KOD_RVBU_95_label),
      color = threshold_colors[2],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    annotate("text", 
    x = pmin(pmax(KOD_RVBU_99, 0.05), 0.95) + 0.02, 
    y = 0.5, 
    label = paste("99%:", KOD_RVBU_99_label), 
    color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) +
    labs(
      x = "BirdNET Confidence Scores",
      y = "Prediction (0/1) "
    ) +
    theme_minimal() +
    theme(text = element_text(family = "A"),
      axis.title = element_text(family = "A", size = 16, face = 'bold'),
      axis.text = element_text(family = "A", size = 12)
    ) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))
```

##SPECIES- GRCO
#Reading the validation tables
```{r}
#Guwahati
GUW_GRCO_annotations <- read.delim ('D:/COVID-Soundscapes/Guwahati/GUW_validation/GRCO/GUW_GRCO_validation_table.txt', header = TRUE, sep = "\t")

#Delhi
DEL_GRCO_annotations <- read.delim ('D:/COVID-Soundscapes/Delhi/DEL_validation/GRCO/DEL_GRCO_validation_table.txt', header = TRUE, sep = "\t")

#Kodaikanal
KOD_GRCO_annotations <- read.delim ('D:/COVID-Soundscapes/Kodaikanal/KOD_validation/GRCO/KOD_GRCO_validation_table.txt', header = TRUE, sep = "\t")
```

#Extracting the confidence score values and adding it as a separate column
```{r}
#Guwahati
GUW_GRCO_annotations$conf_score <- substr(GUW_GRCO_annotations$Begin.File, 1, 5)

#Delhi
DEL_GRCO_annotations$conf_score <- substr(DEL_GRCO_annotations$Begin.File, 1, 5)

#Kodaikanal
KOD_GRCO_annotations$conf_score <- substr(KOD_GRCO_annotations$Begin.File, 1, 5)
```

#Fitting a logistic regression model and plotting the relationship between validation and confidence scores
```{r}
windowsFonts(A = windowsFont("Century Gothic")) 

#Guwahati
GUW_GRCO_annotations$conf_score <- as.numeric (GUW_GRCO_annotations$conf_score) 

GUW_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = GUW_GRCO_annotations) 

prediction_values <- seq(0, 1, .001) 

predictions <- predict(GUW_GRCO_LR, list(conf_score = prediction_values),  type = 'r') 

GUW_GRCO_90 <- (log(.90 / (1 - .90)) - GUW_GRCO_LR$coefficients[1]) / GUW_GRCO_LR$coefficients[2] #confidence scores beyond 0.262 will yield 90% true detections
GUW_GRCO_95 <- (log(.95 / (1 - .95)) - GUW_GRCO_LR$coefficients[1]) / GUW_GRCO_LR$coefficients[2] #confidence scores beyond 0.332 will yield 95% true detections
GUW_GRCO_99 <- (log(.99 / (1 - .99)) - GUW_RVBU_LR$coefficients[1]) / GUW_RVBU_LR$coefficients[2] #this confidence score is unattainable

GUW_GRCO_90_label <- round(GUW_GRCO_90, 3)
GUW_GRCO_95_label <- round(GUW_GRCO_95, 3)
GUW_GRCO_99_label <- round(GUW_GRCO_99, 3)

GUW_GRCO_LR_plot <- ggplot(GUW_GRCO_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
    geom_vline(xintercept = GUW_GRCO_90, color = threshold_colors[1], size = 1) +
    geom_vline(xintercept = GUW_GRCO_95, color = threshold_colors[2], size = 1) +
    geom_vline(xintercept = GUW_GRCO_99, color = threshold_colors[3], size = 1) +
    annotate("text",
      x = pmin(pmax(GUW_GRCO_90, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("90%:", GUW_GRCO_90_label),
      color = threshold_colors[1],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    annotate("text",
      x = pmin(pmax(GUW_GRCO_95, 0.05), 0.95) + 0.02,
      y = 0.5,
      label = paste("95%:", GUW_GRCO_95_label),
      color = threshold_colors[2],
      family = "A",
      angle = 90,
      hjust = 0.5,
      size = 5
    ) +
    #annotate("text", x = pmin(pmax(GUW_GRCO_99, 0.05), 0.99) + 0.02, y = 0.5, label = paste("99%:", GUW_GRCO_99_label),  color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) +
  labs(x = "BirdNET Confidence Scores",y = "Prediction (0/1)") + 
  theme_minimal() + theme(text = element_text(family = "A"), axis.title = element_text(family = "A", size = 16, face = 'bold'), axis.text = element_text(family = "A", size = 12)) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))

#Delhi
DEL_GRCO_annotations$conf_score <- as.numeric (DEL_GRCO_annotations$conf_score) 

DEL_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = DEL_GRCO_annotations) 

prediction_values <- seq(0, 1, .001) 

predictions <- predict(DEL_GRCO_LR, list(conf_score = prediction_values),  type = 'r')

#confiscore score threshold is unattainable as there are too few detections.

DEL_GRCO_LR_plot <- ggplot(DEL_GRCO_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
    #geom_vline(xintercept = DEL_GRCO_90, color = threshold_colors[1], size = 1) +
    #geom_vline(xintercept = DEL_GRCO_95, color = threshold_colors[2], size = 1) +
    #geom_vline(xintercept = DEL_GRCO_99, color = threshold_colors[3], size = 1) +
    #annotate("text", x = pmin(pmax(DEL_GRCO_90, 0.05), 0.95) + 0.02, y = 0.5, label = paste("90%:", DEL_GRCO_90_label), color = threshold_colors[1], family = "A", angle = 90, hjust = 0.5, size = 5) +
    #annotate("text", x = pmin(pmax(DEL_GRCO_95, 0.05), 0.95) + 0.02, y = 0.5, label = paste("95%:", DEL_GRCO_95_label), color = threshold_colors[2], family = "A", angle = 90, hjust = 0.5, size = 5) +
    #annotate("text", x = pmin(pmax(GUW_GRCO_99, 0.05), 0.99) + 0.02, y = 0.5, label = paste("99%:", GUW_GRCO_99_label),  color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) +
  labs(x = "BirdNET Confidence Scores",y = "Prediction (0/1)") + 
  theme_minimal() + theme(text = element_text(family = "A"), axis.title = element_text(family = "A", size = 16, face = 'bold'), axis.text = element_text(family = "A", size = 12)) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))

#Kodaikanal
KOD_GRCO_annotations$conf_score <- as.numeric (KOD_GRCO_annotations$conf_score) 

KOD_GRCO_LR <- glm(valid ~ conf_score, family = 'binomial', data = KOD_GRCO_annotations) 

prediction_values <- seq(0, 1, .001) 

predictions <- predict(KOD_GRCO_LR, list(conf_score = prediction_values),  type = 'r') 

#confiscore score threshold is unattainable as there are too few detections.

KOD_GRCO_LR_plot <- ggplot(KOD_GRCO_annotations, aes(x = conf_score, y = valid)) +
    geom_point(alpha = 0.5, size = 4) +
    geom_line(
      data = data.frame(
        x = prediction_values,
        y = predictions),
      aes(x = x, y = y),
      color = curve_color,
      linewidth = 1.2
    ) +
    #geom_vline(xintercept = KOD_GRCO_90, color = threshold_colors[1], size = 1) +
    #geom_vline(xintercept = KOD_GRCO_95, color = threshold_colors[2], size = 1) +
    #geom_vline(xintercept = KOD_GRCO_99, color = threshold_colors[3], size = 1) +
    #annotate("text", x = pmin(pmax(KOD_GRCO_90, 0.05), 0.95) + 0.02, y = 0.5, label = paste("90%:", KOD_GRCO_90_label), color = threshold_colors[1], family = "A", angle = 90, hjust = 0.5, size = 5) +
    #annotate("text", x = pmin(pmax(KOD_GRCO_95, 0.05), 0.95) + 0.02, y = 0.5, label = paste("95%:", KOD_GRCO_95_label), color = threshold_colors[2], family = "A", angle = 90, hjust = 0.5, size = 5 ) +
    #annotate("text", x = pmin(pmax(KOD_GRCO_99, 0.05), 0.99) + 0.02, y = 0.5, label = paste("99%:", KOD_GRCO_99_label),  color = threshold_colors[3], family = "A", angle = 90, hjust = 0.5, size = 5) +
  labs(x = "BirdNET Confidence Scores",y = "Prediction (0/1)") + 
  theme_minimal() + theme(text = element_text(family = "A"), axis.title = element_text(family = "A", size = 16, face = 'bold'), axis.text = element_text(family = "A", size = 12)) +
    scale_x_continuous(limits = c(0, 1)) +
    scale_y_continuous(limits = c(0, 1))

#It is not possible to obtain confidence thresholds since there are too few detections.
```

