---
title: "04_vocal_activity"
output: html_document
date: "2025-08-20"
---

---
#This script is to calculate the diurnal vocal activity patterns of each species.

##Loading required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)
library(lubridate)
```

##Reading the BirdNET annotation data
```{r}
#Guwahati:
GUW_annotations <- read.csv ('C:/Users/pavit/OneDrive/Desktop/IISER-Tirupati-FA-2025_App/COVIDSoundscapes-project/BirdNET-vocal-activity/results/GUW_BirdNET_annotations.csv')

GUW_metadata <- read.csv ('E:/CovidSoundscape project/Acoustics data/Guwahati/GUW_metadata.csv')
```

##Calculating  sampling effort by time of day
```{r}
#Guwahati
GUW_effort <- GUW_annotations %>%
  dplyr::select(site_code, date, time) %>%
  distinct() %>%
  arrange(time) %>%
  count(time) %>%
  rename(., effort = n)
```

##Filtering the detections of our species of interest whose confidence scores are more than the calibrated species- and location specific thresholds
```{r}
#Guwahati: 
##SPECIES- RVBU
GUW_RVBU <- GUW_annotations %>%
  dplyr::filter (common_name == 'Red-vented Bulbul') %>%
  dplyr::filter (confidence >= 0.5)
  
  dplyr::filter (confidence >= 0.625)
```

##Calculating vocal activity as the percent detections at each hour of the day
```{r}
#Guwahati: 
##SPECIES- RVBU
GUW_RVBU_vocal_activity <- GUW_RVBU %>%
  group_by(time) %>%
  count (.) %>%
  rename (detections = n) %>%
  left_join(., GUW_effort, by = "time") %>% #adding the column of sampling effort for each hour of the day
  mutate (normalized_detections = detections/effort) %>% #normalizing detections by dividing it by effort
  ungroup () %>%
  mutate(total_normalized_detections = sum(normalized_detections)) %>% 
  mutate(percent_normalized_detections = (normalized_detections / total_normalized_detections) * 100) #this gives us the percent vocal activity for each hour of the day
```






