---
title: "04_vocal_activity"
output: html_document
date: "2025-08-20"
---

---
#This script is to calculate the diurnal vocal activity patterns of each species.

##Loading required libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(sf)
library(raster)
library(lubridate)
```

##Reading the BirdNET annotation data
```{r}
#Guwahati:
GUW_annotations <- read.csv ('C:/Users/pavit/OneDrive/Desktop/IISER-Tirupati-FA-2025_App/COVIDSoundscapes-project/BirdNET-vocal-activity/results/GUW_BirdNET_annotations.csv')
```

##Calculating  sampling effort by time of day
```{r}
#Guwahati
GUW_effort <- GUW_annotations %>%
  dplyr::select(site_code, date, time) %>%
  distinct() %>%
  arrange(time) %>%
  count(time) %>%
  rename(., effort = n)
```

##Filtering the detections of our species of interest whose confidence scores are more than the calibrated species- and location specific thresholds
```{r}
#Guwahati: 
##SPECIES- RVBU
GUW_RVBU <- GUW_annotations %>%
  dplyr::filter (common_name == 'Red-vented Bulbul') %>%
  dplyr::filter (confidence >= 0.596)

##SPECIES- GRCO
GUW_GRCO <- GUW_annotations %>%
  dplyr::filter (common_name == 'Greater Coucal') %>%
  dplyr::filter (confidence >= 0.332)
```

##Calculating vocal activity as the percent detections at each hour of the day
```{r}
#Guwahati: 
##SPECIES- RVBU
GUW_RVBU_vocal_activity <- GUW_RVBU %>%
  group_by(time) %>%
  count (.) %>%
  rename (detections = n) %>%
  left_join(., GUW_effort, by = "time") %>% #adding the column of sampling effort for each hour of the day
  mutate (normalized_detections = detections/effort) %>% #normalizing detections by dividing it by effort
  ungroup () %>%
  mutate(total_normalized_detections = sum(normalized_detections)) %>% 
  mutate(percent_normalized_detections = (normalized_detections / total_normalized_detections) * 100) %>% #this gives us the percent vocal activity for each hour of the day
  mutate (species = 'RVBU')
  
##SPECIES- GRCO
GUW_GRCO_vocal_activity <- GUW_GRCO %>%
  group_by(time) %>%
  count (.) %>%
  rename (detections = n) %>%
  left_join(., GUW_effort, by = "time") %>% 
  mutate (normalized_detections = detections/effort) %>% 
  ungroup () %>%
  mutate(total_normalized_detections = sum(normalized_detections)) %>% 
  mutate(percent_normalized_detections = (normalized_detections / total_normalized_detections) * 100) %>%
  mutate (species = 'GRCO')
```

##Plotting the diel vocal activity patterns
```{r}
#Guwahati: 

GUW_vocal_activity <- bind_rows (GUW_GRCO_vocal_activity, GUW_RVBU_vocal_activity) %>%
  filter (!time %in% c('43000','183000')) %>%
  ggplot(., aes(x= time, y= percent_normalized_detections, group = species)) +
  geom_line (linewidth = 0.2) +
  facet_wrap (~ species, nrow = 2) +
  theme_bw() +
    theme(text = element_text(family = "A"),
      axis.title = element_text(family = "A", size = 16, face = 'bold'),
      axis.text = element_text(family = "A", size = 12)) +
  labs(x = 'Time of day', y = 'Vocal activity (Percent normalized detections)') 

GUW_vocal_activity + theme(strip.background = element_blank(), strip.text.x = element_blank(), strip.text.y = element_blank() )
  


```






